---
title: "covid 19"
author: "Emilee Coronel"
date: "10/5/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata$state<-as.factor(kRdata$state)
kRdata$state<-as.factor(kRdata$result)
kRdata$sex<-as.factor(kRdata$sex)
kRdata$sex<-as.factor(kRdata$sex)
```



### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this


```{r}
barchartGC(~result + Sex,data=kRdata)
```


From this bar chart, we can see that more women are alive than men, with woren being more prominent than men. The amount of women and
the men that died were not substantial. All and all, there were more men that died than women.  






```{r}
barchartGC(~class + Sex,data=ucdavis1, type="percent")

```


We can see from the second bar chart that about 55-56% of women were alive vs about 40* of men,and from the dead category we can see more men died vs the amount of women, From both bar charts We can see that women did better in this study.  






  
## Numerical results


```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```


We now have the correct precentages, whereos we can see that the alive group was F 56N and M 44%.In the dead group F ranged in at 36% and M came in at 64%. So we can see here there is a backward relationship between males and females, Females were prodominalty alive wheras men were prodominatley dead, 1.4 F ended up dead, where 3% of males ended up dead. Males deaths were double the rate of women.  



## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
```





We now have the correct percentages, where we can see that the alive group was F 56N and M 44%.In the dead group F ranged in at 36% and M came in at 64%. So we can see here there is a backward relationship between males and females, Females were predominantly alive whereas men were predominantly dead, 1.4 F ended up dead, where 3% of males ended up dead. Males deaths were double the rate of women.  



```{r}
fisher.test(table1)
```


The fisher test give us the odds of dying for males ts (44/1411) and the odds of the females dying is (25/1831), when you divide these two numbers you get 2,3. The odds ratio shows that males are 2.3 times more likely to die. I also see the confidence intervals are 1,4 to 3,9. 



## Conclusions


 In conclusion we can see a big statistical difference between men and women. Seemingly more men die of covid vs women. Our p-value is very small, so therefore we reject the null hypothesis.More importantly we can see more men die from this disease vs women.  




